Binocular integration of chromatic and luminance signals

### 2024-04-19 manuscript.tex
Various edits and fixes

Much progress has been made in understanding how the brain combines
signals from the two eyes. However, most of this work has involved
achromatic (black and white) stimuli, and it is not clear if the same
processes apply in colour-sensitive pathways. In our first experiment,
we measured contrast discrimination (`dipper') functions for four key
ocular configurations (monocular, binocular, half-binocular and
dichoptic), for achromatic, isoluminant red/green and isoluminant
blue/yellow sine-wave grating stimuli. We find a similar pattern of
results across stimuli, implying equivalently strong interocular
suppression within each pathway. Our second experiment measured
dichoptic masking within and between pathways using the method of
constant stimuli. Masking was strongest within-pathway, and weakest
between chromatic and achromatic mechanisms. Finally, we repeated the
dipper experiment using temporal luminance modulations, which produced
slightly weaker interocular suppression than for spatially modulated
stimuli. We interpret our results in the context of a contemporary
two-stage model of binocular contrast gain control, implemented here
using a hierarchical Bayesian framework. Posterior distributions of the
weight of interocular suppression overlapped with a value of 1 for all
dipper data sets, and the model captured well the pattern of threshold
and slope values from all three experiments.

@@

The detailed pattern of thresholds across these four conditions is
complex, and for achromatic stimuli has several distinctive features
that have been replicated in multiple studies. At low pedestal
contrasts, the binocular condition yields lower thresholds than the
monocular condition; this result is attributed to physiological
binocular summation by neurons responsive to signals from both eyes
(Baker et al., 2018; Campbell \& Green, 1965). However, at high pedestal
contrasts the `handle' regions of the dipper functions for these
conditions converge: a consequence of interocular suppression
compensating for the increased excitation during binocular stimulation
(Legge, 1984; Maehara \& Goryo, 2005). The half-binocular condition
avoids confounding the number of eyes seeing the target with the number
of eyes seeing the pedestal (Meese et al., 2006). The pedestal is always
binocular in this condition, whereas the target increment is monocular,
and thresholds are consistently higher than in the binocular condition
across the full range of pedestal contrasts. This demonstrates that
binocular summation occurs across the full contrast range, when the
pedestal ocularity is appropriately controlled. Finally, the dichoptic
condition produces extremely strong masking of the target, such that
when the pedestal is visible, the target must equal or exceed its
contrast to be detectable (Baker \& Meese, 2007; Legge, 1979; Maehara \&
Goryo, 2005). The characteristic pattern of dipper functions is
well-described by the gain control model of Meese et al. (2006) for
achromatic stimuli.

@@

There has not yet been a detailed investigation of binocular contrast
discrimination in either of these chromatic pathways, however there is
reason to believe they may differ from the achromatic pathway. At
detection threshold, binocular summation is greater for chromatic versus
achromatic stimuli (Simmons, 2005), implying a more linear initial stage
of processing. For cross-orientation masking, there are differences in
the magnitude of masking between chromatic (red/green) and achromatic
stimuli (Kim et al., 2013; Medina \& Mullen, 2009), as well as
differences in their temporal dynamics (Kim \& Mullen, 2015). There are
also interactions between chromatic and achromatic pathways both within
(Chen et al., 2000) and between (Kingdom \& Libenson, 2015; Mullen et
al., 2014) the eyes, yet these have not been fully explored for
arrangements where the target and mask have the same orientation.
Finally, the neurophysiological underpinnings of colour vision are
distinct from those of the achromatic system. In primary visual cortex
(V1), chromatic signals are processed in `blob' regions that are
revealed by cytochrome oxidase staining (Horton \& Hubel, 1981). The
blob regions appear to be largely monocular (Livingstone \& Hubel,
1984), suggesting that binocular combination for chromatic stimuli might
occur later than for achromatic stimuli, and perhaps be subject to
different constraints.

A further property of `blob' regions is that they are less strongly
orientation-tuned than other regions of V1 (Horton \& Hubel, 1981;
Livingstone \& Hubel, 1984), and biased towards low spatial frequencies
(Edwards et al., 1995; Tootell et al., 1988; though see Economides et
al., 2011). Our recent work (Segala et al., 2023) has investigated
binocular combination for flickering discs of luminance, which are
DC-balanced across time (i.e.~the time-averaged luminance is equal to
the background). Steady-state EEG responses from early visual cortex and
psychophysical contrast matching data were both consistent with weak
interocular suppression when using this stimulus arrangement. This is
very different from the `ocularity invariance' (where binocular and
monocular stimuli appear equal) that is well-established when using
DC-balanced periodic stimuli such as sine-wave gratings, and implies
strong interocular suppression (Baker \& Wade, 2017; Meese et al., 2006;
Moradi \& Heeger, 2009). There is additional psychophysical evidence
from matching studies that binocular combination can be close to linear
for luminance increments (Anstis \& Ho, 1998; Levelt, 1965),
particularly against a dark background (Baker et al., 2012). Both
structural and functional data therefore imply that binocular
combination may differ between spatial and temporal contrast.

@@

All stimuli were presented on an Iiyama VisionMaster Pro 510 CRT
monitor, with a refresh rate of 100Hz, and a resolution of 1024 x 768
pixels. The display was driven by a ViSaGe MkII stimulus generator
(Cambridge Research Systems Ltd., Kent, UK) running in 42-bit colour
mode (14 bits per colour channel). We presented stimuli to the left and
right eyes independently using a four-mirror stereoscope with
front-silvered mirrors. The display was luminance calibrated using a
ColourCal photometer (Cambridge Research Systems), and gamma corrected
by fitting a four-parameter gamma function to the output of each CRT
gun. The maximum luminance was 87 cd/m\(^2\). We also measured the
spectral output of each phosphor using a Jaz spectroradiometer (Ocean
Insight, Florida), and used these measurements to convert between LMS
(cone) space and the monitor RGB coordinates. We express contrast values
in normalized units, relative to the monocular threshold for each
stimulus type.

@@

Before beginning primary data collection, each participant in
Experiments 1 and 2 completed an isoluminance adjustment task. Grating
stimuli were presented that counterphase flickered at 5Hz, with contrast
variations defined about either the L-M or S-(L+M) plane in cone space.
Participants used a trackball to dynamically adjust the colour angle of
the stimulus to minimise the percept of flicker. Each participant
completed ten such trials for each colour plane, and the average angle
across repetition was taken as the isoluminant point, and used to
generate stimuli for the main experiment for that participant. Settings
were very similar across participants for the S-(L+M) direction, and
somewhat more heterogeneous for the L-M direction (see
Figure~\ref{fig-isofig}).

@@

In Experiment 1, participants completed a two-interval-forced-choice
(2IFC) contrast discrimination task. Stimuli were presented for 200ms,
with an interstimulus interval of 400ms. Each interval was indicated by
an auditory beep, and participants made their responses using a
two-button trackball. Correct responses were followed by a high-pitched
tone, and incorrect responses by a low-pitched tone. Each block of the
experiment tested a single pedestal contrast level and lasted around 12
minutes. On each trial the target contrast level was determined by a
3-down-1-up staircase procedure. There were 8 interleaved staircases in
total; four stimulus arrangements (see Figure~\ref{fig-exampledips})
combined factorially with two target eye assignments. Each pedestal
contrast was repeated 3 times by each participant, and the block order
was randomised. The experiment lasted around 4 hours per participant for
each chromatic condition, and took place over the course of several
weeks. In total, the experiment consisted of 67978 trials (pooled across
participants).

@@

where \(C_L\) and \(C_R\) are the (normalized) contrasts displayed to
the left and right eyes, and \(m\), \(S\), \(\omega\), \(p\), \(q\) and
\(Z\) are free parameters in the model. A further free parameter, \(k\),
represents additive internal noise, and is used to convert the model
outputs to either d-prime or threshold values (note that in the original
model specification (Meese et al., 2006) this parameter was called
\(\sigma\), but we use the \(k\) symbol here to avoid confusion with the
standard deviation of the cumulative Gaussian used when fitting the
psychometric functions to estimate thresholds). Thresholds are defined
by iteratively adjusting the target contrast until the following
equality is satisfied:

@@

\begin{equation}\protect\hypertarget{eq-dprime}{}{d' = \frac{\mathrm{Stage2}_{\mathrm{target+pedestal}} - \mathrm{Stage2}_{\mathrm{pedestal}}}{k/0.954},}\label{eq-dprime}\end{equation}

where the constant 0.954 reflects the value of \(d'\) at the 75\%
correct point on the psychometric function used to estimate threshold.

@@

where \emph{AC} represents the achromatic contrast, \emph{RG} represents
the red/green contrast, \emph{BY} represents the blue/yellow contrast,
and \(\omega_A\), \(\omega_R\) and \(\omega_B\) are the accompanying
weights of interocular suppression. There is an equivalent expression
for the right eye, and for each of the two isoluminant chromatic
pathways. To simplify the model, and avoid free parameters that are
poorly constrained by the data, we fixed several parameters (\emph{p},
\emph{q}, \emph{m}, \emph{S}, \emph{k}, and \(\omega\) for the
within-pathway suppression) at the values from the fits from Experiment
1 (Table~\ref{tbl-parametertable}, lower rows). This left 9 free
parameters: a \(Z\) parameter for each mechanism, and six
cross-mechanism weights of interocular suppression. These parameters
were again estimated within a Bayesian hierarchical framework, using the
trial-wise data from Experiment 2. Note that the model as specified does
not currently include monocular suppression between different pathways,
as we did not collect any data for these conditions. Previous work (Chen
et al., 2000) has measured such interactions, and they could in
principle be incorporated into the denominator of either stage 1 or
stage 2 in the model.

@@

Following Meese et al. (2006), we also inspected the slope of the
psychometric function for each condition (see
Figure~\ref{fig-dipperfig}d-f). At detection threshold, slopes were
relatively steep for all stimuli, with values around \(\beta=4\). As
pedestal contrasts increased, slopes linearized and reduced to around
\(\beta=1.3\) (Foley \& Legge, 1981; Meese et al., 2006), and remained
shallow at high pedestal contrasts. The exception to this was the
dichoptic condition, where slopes became extremely steep at high
dichoptic mask contrasts, consistent with previous observations (Baker
et al., 2013; Meese et al., 2006). This was clear for all three data
sets, with slope values in the range \(4 < \beta < 8\). However, we
observe that slope estimates are more variable than threshold estimates
(note the large error bars), particularly when using adaptive
staircases, which deploy most trials close to threshold. Our second
experiment therefore investigated the slope of the psychometric function
in more detail for dichoptic masking using the method of constant
stimuli, as well as exploring dichoptic interactions between chromatic
channels.

@@

\caption{\label{fig-MCSfig}Summary of data from Experiment 2. Panels
(a-c) show psychometric functions for each condition, pooled across
participants (600 trials per target contrast level). Panels (d-f) show
threshold and slope estimates for individual participants (points) and
the boundary of the posterior density estimates for fits to the pooled
data (ellipses). Panel (g) shows the average threshold elevation factor
for each combination of target and mask stimulus. Panel (h) shows the
geometric mean psychometric slope value for each masking condition in
Weibull \(\beta\) units. Contrast values are expressed relative to the
target only threshold for each stimulus type.}

@@

We additionally implemented a hierarchical Bayesian version of the model
to estimate full posterior parameter distributions. This model was fit
simultaneously to the full trial-by-trial data from all participants who
participated in a given experiment (i.e.~the model was fitted separately
to each of the four dipper data sets). Figure~\ref{fig-bayesianmodel}a-d
summarises the model behaviour, which displays the same pattern of
dipper functions as we found empirically. Modal posterior parameter
values (maximum a posteriori (MAP) estimates) are given in the lower
rows of Table~\ref{tbl-parametertable}. These correspond quite closely
to the parameters from the simplex fitting and the original Meese et al.
(2006) parameters (reproduced in the first row of the table for
reference). The panels in the right and upper margins of
Figure~\ref{fig-bayesianmodel} show posterior distributions for each
model parameter. Note in particular that the posterior distribution for
the weight of interocular suppression (\(\omega\), top right panel)
overlaps 1 for each experiment, consistent with the strong dichoptic
masking observed in the threshold data. We note that the distribution
for \(\omega\) in the flickering disc condition is somewhat lower than
for the other three data sets. However this difference is not meaningful
according to widely accepted criteria such as comparing the 95\%
intervals of the distribution to a value of \(\omega=1\).

@@

Finally we fitted an extended model, that included interocular
suppression between the different pathways, to the data of Experiment 2.
Figure~\ref{fig-MCSmodel} shows the model curves (panels a-c), which
correspond closely to the data in Figure~\ref{fig-MCSfig}. One striking
discrepancy is that the model predicts a region of negative d-prime for
the case where the target and dichoptic mask have the same chromaticity
(see the curved regions below 50\% correct). This happens in the model
because low contrast targets suppress the mask more than they excite the
detecting mechanism, producing a net decrease in response. The feature
(termed a `swan function') was not generally present in our empirical
data, though it can be observed for one participant in
Figure~\ref{fig-individualMCS}b,c.~Our previous work (Baker et al.,
2013) has found evidence for this phenomenon, but it generally requires
very high mask contrasts to be measurable empirically, and there may
also be some individual differences; both factors might explain its
absence here.

Figure~\ref{fig-MCSmodel}d,e shows the model estimates for the weight of
interocular suppression in each combination of target and mask
chromaticity. The weakest suppression was between achromatic targets and
masks and blue/yellow targets and masks, and the strongest suppression
was between red/green targets and blue/yellow masks. Note that the
weight parameter value of 1.21 for this condition is higher than that of
the within-channel weights (values around 1), even though the
within-channel conditions generate more threshold elevation. In the
model, high thresholds for within-channel dichoptic masks occur because
the target and mask are summed together, which causes additional masking
(see Baker et al., 2013). There do not appear to be any other salient
patterns in the suppressive weight estimates.

@@

\caption{\label{fig-MCSmodel}Summary of the model fit to the data of
Experiment 2. Panels (a-c) show model psychometric functions in the same
format as those in Figure 5. Panel (d) shows the fitted suppressive
weights (\(\omega\) values) estiamted by the model, and panel (e) shows
the posterior distributions for each weight parameter. The lower right
plot gives the legends for the upper row (top) and the posterior
distributions in panel (e).}

@@

\begin{itemize}
\tightlist
\item
  Binocular combination of isoluminant chromatic stimuli is similar to
  that for achromatic stimuli
\item
  Interocular suppression is strongest within a post-retinal pathway,
  and weakest between achromatic and chromatic pathways
\item
  Binocular combination occurs similarly for spatial and temporal
  luminance modulations
\end{itemize}

@@

The weight of interocular suppression is represented by the model
parameter, \(\omega\) (Equation~\ref{eq-s1l} and Equation~\ref{eq-s1r}).
Posterior distributions of this parameter overlapped 1 for all of our
dipper function data sets (see top right panel in
Figure~\ref{fig-bayesianmodel}), indicating strong interocular
suppression regardless of post-retinal pathway. The flickering disc
stimuli produced a slightly lower suppression estimate, with
\(\omega \sim 0.9\). This may indicate slightly weaker suppression
between the eyes for temporal modulations, but it is much less extreme
than our recent estimates using EEG and matching paradigms (Segala et
al., 2023). One difference between these studies is the luminance of the
background, which was set to black for the experiments of Segala et al.
(2023), but here was set at the mean luminance. Other studies using
static luminance increments have also reported differences in the
character of binocular combination that are attributable to background
luminance (Baker et al., 2012), so this may explain the differences
between studies. We also note that the weaker interocular suppression
for flickering disc stimuli appears to be largely due to one participant
(see Figure~\ref{fig-individualdiscs}c), so individual differences might
also play a role. Future work should manipulate the background luminance
systematically to better understand how this modulates interocular
suppression.

@@

In V1, chromatic stimuli are processed in `blob' regions that are
largely monocular as they fall within ocular dominance columns
(Livingstone \& Hubel, 1984). Since our psychophysical results indicate
that interocular suppression is equally strong within chromatic and
achromatic pathways, it may be that blob regions include interocular
suppression. Presumably binocular summation of signals occurs at a later
stage of processing for chromatic stimuli, but this does not lead to
different perceptual outcomes. This physiological arrangement might also
be responsible for the pattern of between-pathway interocular
suppression (see Figure~\ref{fig-MCSfig}g). If blobs process signals
from both the L-M and S-(L+M) pathways, their close physical proximity
might explain the stronger suppression between the two chromatic
pathways than between the chromatic and (more distant) achromatic
pathways.

@@

@@

>>>>

Much progress has been made in understanding how the brain combines
signals from the two eyes. However, most of this work has involved
achromatic (black and white) stimuli, and it is not clear if the same
processes apply in colour-sensitive pathways. In our first experiment,
we measured contrast discrimination (`dipper') functions for four key
ocular configurations (monocular, binocular, half-binocular and
dichoptic), for achromatic, isoluminant red/green and isoluminant
blue/yellow sine-wave grating stimuli. We find a similar pattern of
results across stimuli, implying equivalently strong interocular
suppression within each pathway. Our second experiment measured
dichoptic masking within and between pathways using the method of
constant stimuli. Masking was strongest within-pathway, and weakest
between blue/yellow and achromatic mechanisms. Finally, we repeated the
dipper experiment using temporal luminance modulations, which produced
slightly weaker interocular suppression than for spatially modulated
stimuli. We interpret our results in the context of a contemporary
two-stage model of binocular contrast gain control, implemented here
using a hierarchical Bayesian framework. Posterior distributions of the
weight of interocular suppression overlapped with a value of 1 for all
dipper data sets, and the model captured well the pattern of thresholds
from all three experiments.

@@

The detailed pattern of thresholds across these four conditions is
complex, and for achromatic stimuli has several distinctive features
that have been replicated in multiple studies. At low pedestal
contrasts, the binocular condition yields lower thresholds than the
monocular condition; this result is attributed to physiological
binocular summation by neurons responsive to signals from both eyes
(Baker et al., 2018; Campbell \& Green, 1965). However, at high pedestal
contrasts the `handle' regions of the dipper functions for these
conditions converge: a consequence of interocular suppression
compensating for the increased excitation during binocular stimulation
(Legge, 1984; Maehara \& Goryo, 2005). The half-binocular condition
avoids confounding the number of eyes seeing the target with the number
of eyes seeing the pedestal (Meese et al., 2006). The pedestal is always
binocular in this condition, whereas the target increment is monocular,
and thresholds are consistently higher than in the binocular condition
across the full range of pedestal contrasts. This demonstrates that
binocular summation occurs across the full contrast range, when the
pedestal ocularity is appropriately controlled. Finally, the dichoptic
condition produces extremely strong masking of the target, such that
when the pedestal is visible, the target must equal or exceed its
contrast to be detectable (Baker \& Meese, 2007; Legge, 1979; Maehara \&
Goryo, 2005). The characteristic pattern of dipper functions (right
panel of Figure~\ref{fig-exampledips}) is well-described by the gain
control model of Meese et al. (2006) for achromatic stimuli.

@@

There has not yet been a detailed investigation of binocular contrast
discrimination in either of these chromatic pathways, however there is
reason to believe they may differ from the achromatic pathway. At
detection threshold, binocular summation is greater for chromatic versus
achromatic stimuli (Simmons, 2005), implying a more linear initial stage
of processing. For cross-orientation masking, there are differences in
the magnitude of masking between chromatic (red/green) and achromatic
stimuli (Kim et al., 2013; Medina \& Mullen, 2009), as well as
differences in their temporal dynamics (Kim \& Mullen, 2015). There are
also interactions between chromatic and achromatic pathways both within
(Chen et al., 2000) and between (Kingdom \& Libenson, 2015; Mullen et
al., 2014) the eyes, yet these have not been fully explored for
arrangements where the target and mask have the same orientation.
Finally, the neurophysiological underpinnings of colour vision are
distinct from those of the achromatic system. In primary visual cortex
(V1), chromatic signals are processed in `blob' regions that are
revealed by cytochrome oxidase staining (Horton \& Hubel, 1981). The
blob regions appear to be largely monocular (Livingstone \& Hubel,
1984), suggesting that binocular combination for chromatic stimuli might
occur separately than for achromatic stimuli, and perhaps be subject to
different constraints.

A further property of `blob' regions is that they are less strongly
orientation-tuned than other regions of V1 (Horton \& Hubel, 1981;
Livingstone \& Hubel, 1984), and biased towards low spatial frequencies
(Edwards et al., 1995; Tootell et al., 1988; though see Economides et
al., 2011). Our recent work (Segala et al., 2023) has investigated
binocular combination for flickering discs of luminance, which are
DC-balanced across time (i.e.~the time-averaged luminance is equal to
the background). Steady-state EEG responses from early visual cortex and
psychophysical contrast matching data were both consistent with weak
interocular suppression when using this stimulus arrangement. This is
very different from the `ocularity invariance' (where binocular and
monocular stimuli appear equal) that is well-established when using
spatially modulated stimuli such as sine-wave gratings, and implies
strong interocular suppression (Baker \& Wade, 2017; Meese et al., 2006;
Moradi \& Heeger, 2009). There is additional psychophysical evidence
from matching studies that binocular combination can be close to linear
for luminance increments (Anstis \& Ho, 1998; Levelt, 1965),
particularly against a dark background (Baker et al., 2012). Both
structural and functional data therefore imply that binocular
combination may differ between spatial and temporal contrast.

@@

All stimuli were presented on an Iiyama VisionMaster Pro 510 CRT
monitor, with a refresh rate of 100Hz, and a resolution of 1024 x 768
pixels. The display was driven by a ViSaGe MkII stimulus generator
(Cambridge Research Systems Ltd., Kent, UK) running in 42-bit colour
mode (14 bits per colour channel). We presented stimuli to the left and
right eyes independently using a four-mirror stereoscope with
front-silvered mirrors. The display was luminance calibrated using a
ColourCal photometer (Cambridge Research Systems), and gamma corrected
by fitting a four-parameter gamma function to the output of each CRT
gun. The maximum luminance was 87 cd/m\(^2\). We also measured the
spectral output of each phosphor using a Jaz spectroradiometer (Ocean
Insight, Florida), and used these measurements to convert between LMS
(cone) space and the monitor RGB coordinates. We express threshold and
pedestal contrasts in normalized units, relative to the monocular
detection threshold for each stimulus type. A threshold value of 2
therefore means that, relative to monocular detection, twice as much
contrast was required to reach the same level of performance.

@@

Before beginning primary data collection, each participant in
Experiments 1 and 2 completed an isoluminance adjustment task. Grating
stimuli were presented that counterphase flickered at 5Hz, with contrast
variations defined about either the L-M or S-(L+M) plane in cone space.
Participants used a trackball to dynamically adjust the colour angle of
the stimulus to minimise the percept of flicker. Each participant
completed ten such trials for each colour plane, and the average angle
across repetition was taken as the isoluminant point, and used to
generate stimuli in the main experiment for that participant. Settings
were very similar across participants for the S-(L+M) direction, and
somewhat more heterogeneous for the L-M direction (see
Figure~\ref{fig-isofig}).

@@

In Experiment 1, participants completed a two-interval-forced-choice
(2IFC) contrast discrimination task. Stimuli were presented for 200ms,
with an interstimulus interval of 400ms. Each interval was indicated by
an auditory beep, and participants made their responses using a
two-button trackball. Correct responses were followed by a high-pitched
tone, and incorrect responses by a low-pitched tone. Each block of the
experiment tested a single pedestal contrast level and lasted around 12
minutes. The pedestal contrasts were 0, 0.5, 1, 2, 4, 8, 16 and 32\%
Michelson contrast for the achromatic stimulus, and 0, 1, 2, 4, 8, 16,
32 and 64\% of the maximum available contrast for each chromatic
stimulus. On each trial the target contrast level was determined by a
3-down-1-up staircase procedure, moving in logarithmic (3dB) steps.
There were 8 interleaved staircases in total; four stimulus arrangements
(see Figure~\ref{fig-exampledips}) combined factorially with two target
eye assignments. Each pedestal contrast was repeated 3 times by each
participant, and the block order was randomised. The experiment lasted
around 4 hours per participant for each chromatic condition, and took
place over the course of several weeks. In total, the experiment
consisted of 67978 trials (pooled across participants).

@@

where \(C_L\) and \(C_R\) are the (normalized) contrasts displayed to
the left and right eyes, and \(m\), \(S\), \(\omega\), \(p\), \(q\) and
\(Z\) are free parameters in the model. A further free parameter, \(k\),
is used to convert the model outputs to either d-prime or threshold
values (note that in the original model specification (Meese et al.,
2006) this parameter was called \(\sigma\), but we use the \(k\) symbol
here to avoid confusion with the standard deviation of the cumulative
Gaussian used when fitting the psychometric functions to estimate
thresholds). Thresholds are defined by iteratively adjusting the target
contrast until the following equality is satisfied:

@@

\begin{equation}\protect\hypertarget{eq-dprime}{}{d' = \frac{\mathrm{Stage2}_{\mathrm{target+pedestal}} - \mathrm{Stage2}_{\mathrm{pedestal}}}{k/\tau},}\label{eq-dprime}\end{equation}

where the parameter \(\tau\) reflects the value of \(d'\) at detection
threshold. Defining threshold as the 75\% correct point on the
psychometric function (as here) yields a value of
\(\tau = \Phi^{-1}(0.75)\sqrt{2}\) = 0.954 (where \(\Phi^{-1}\) is the
inverse cumulative normal density function). The combined denominator
term (\(k/\tau\)) therefore represents internal additive noise in the
model.

@@

where \emph{AC} represents the achromatic contrast, \emph{RG} represents
the red/green contrast, \emph{BY} represents the blue/yellow contrast,
and \(\omega_A\), \(\omega_R\) and \(\omega_B\) are the accompanying
weights of interocular suppression. There is an equivalent expression
for the right eye, and for each of the two isoluminant chromatic
pathways. To simplify the model, and avoid free parameters that are
poorly constrained by the data, we fixed several parameters (\emph{p},
\emph{q}, \emph{m}, \emph{S}, \emph{k}, and \(\omega\) for the
within-pathway suppression) at the values from the fits from Experiment
1 (Table~\ref{tbl-parametertable}, lower rows). This left 9 free
parameters: a \(Z\) parameter for each mechanism, and six
cross-mechanism weights of interocular suppression. These parameters
were again estimated within a Bayesian hierarchical framework, using the
binomial proportion correct data from Experiment 2. Note that the model
as specified does not currently include monocular suppression between
different pathways, as we did not collect any data for these conditions.
Previous work (Chen et al., 2000) has measured such interactions, and
they could in principle be incorporated into the denominator of either
stage 1 or stage 2 in the model.

@@

Following Meese et al. (2006), we also inspected the slope of the
psychometric function for each condition (see
Figure~\ref{fig-dipperfig}d-f), as this provides information about the
effective gradient of the underlying contrast response function. At
detection threshold, slopes were relatively steep for all stimuli, with
values around \(\beta=4\). As pedestal contrasts increased, slopes
linearized and reduced to around \(\beta=1.3\) (Foley \& Legge, 1981;
Meese et al., 2006), and remained shallow at high pedestal contrasts.
The exception to this was the dichoptic condition, where slopes became
extremely steep at high dichoptic mask contrasts, consistent with
previous observations (Baker et al., 2013; Meese et al., 2006). This was
clear for all three data sets, with slope values in the range
\(4 < \beta < 8\). However, we observe that slope estimates are more
variable than threshold estimates (note the large error bars),
particularly when using adaptive staircases, which deploy most trials
close to threshold. Our second experiment therefore investigated the
slope of the psychometric function in more detail for dichoptic masking
using the method of constant stimuli, as well as exploring dichoptic
interactions between chromatic channels.

@@

\caption{\label{fig-MCSfig}Summary of data from Experiment 2. Panels
(a-c) show psychometric functions for each condition, pooled across
participants (600 trials per target contrast level). Panels (d-f) show
threshold and slope estimates for individual participants (points) and
the boundary of the posterior density estimates for fits to the pooled
data (ellipses). Panel (g) shows the average threshold elevation factor
for each combination of target and mask stimulus. Panel (h) shows the
geometric mean psychometric slope value for each masking condition,
expressed in Weibull \(\beta\) units. Contrast values are normalized to
the target only threshold for each stimulus type.}

@@

We additionally implemented a hierarchical Bayesian version of the model
to estimate full posterior parameter distributions. This model was fit
simultaneously to the full trial-by-trial data from all participants who
participated in a given experiment (i.e.~the model was fitted separately
to each of the four dipper data sets). Figure~\ref{fig-bayesianmodel}a-d
summarises the model behaviour, which displays the same pattern of
dipper functions as we found empirically. Modal posterior parameter
values (maximum a posteriori (MAP) estimates) are given in the lower
rows of Table~\ref{tbl-parametertable}. These are in a similar range to
the parameters from the simplex fitting and the original Meese et al.
(2006) parameters (reproduced in the first row of the table for
reference). The panels along the right and upper margins of
Figure~\ref{fig-bayesianmodel} show posterior distributions for each
model parameter. Note in particular that the posterior distribution for
the weight of interocular suppression (\(\omega\), top right panel)
overlaps 1 for each experiment, consistent with the strong dichoptic
masking observed in the threshold data. We note that the distribution
for \(\omega\) in the flickering disc condition is somewhat lower than
for the other three data sets. However this difference is not meaningful
according to widely accepted criteria such as comparing the 95\%
intervals of the distribution to a value of \(\omega=1\).

@@

Finally we fitted an extended model, that included interocular
suppression between the different pathways (see
Equation~\ref{eq-s1multi}), to the data of Experiment 2.
Figure~\ref{fig-MCSmodel} shows the model curves (panels a-c), which
correspond closely to the data in Figure~\ref{fig-MCSfig}. One striking
discrepancy is that the model predicts a region of negative d-prime for
the case where the target and dichoptic mask have the same chromaticity
(see the curved regions below 50\% correct). This happens in the model
because low contrast targets suppress the mask more than they excite the
detecting mechanism, producing a net decrease in response. The feature
(termed a `swan function') was not generally present in our empirical
data, though it can be observed for one participant in
Figure~\ref{fig-individualMCS}b,c.~Our previous work (Baker et al.,
2013) has found evidence for this phenomenon, but it generally requires
very high mask contrasts to be measurable empirically, and there may
also be some individual differences; both factors might explain its
absence here.

Figure~\ref{fig-MCSmodel}d,e shows the model estimates for the weight of
interocular suppression in each combination of target and mask
chromaticity. The weakest suppression was between achromatic targets and
masks and blue/yellow targets and masks, and the strongest suppression
was between red/green targets and blue/yellow masks. Note that the
weight parameter value of \(\omega\) = 1.11 for this condition is
slightly higher than that of the within-channel weights (values around
1), even though the within-channel conditions generate more threshold
elevation. In the model, high thresholds for within-channel dichoptic
masks occur because the target and mask are summed together, which
causes additional masking (see Baker et al., 2013). However it is clear
that the posterior distribution of the weight parameter is quite broad
and overlaps 1, so this difference may not be meaningful. There do not
appear to be any other salient patterns in the suppressive weight
estimates.

@@

\caption{\label{fig-MCSmodel}Summary of the model fit to the data of
Experiment 2. Panels (a-c) show model psychometric functions in the same
format as those in Figure 5. Panel (d) shows the fitted suppressive
weights (\(\omega\) values) estimated by the model, and panel (e) shows
the posterior distributions for each weight parameter. The lower right
plot gives the legends for the upper row (top) and the posterior
distributions in panel (e).}

@@

\begin{itemize}
\tightlist
\item
  Binocular combination of isoluminant chromatic stimuli is similar to
  that for achromatic stimuli
\item
  Interocular suppression is strongest within a post-retinal pathway,
  and weakest between the achromatic and blue/yellow pathways
\item
  Binocular combination occurs similarly for spatial and temporal
  luminance modulations
\end{itemize}

@@

The weight of interocular suppression is represented by the model
parameter, \(\omega\) (Equation~\ref{eq-s1l} and Equation~\ref{eq-s1r}).
Posterior distributions of this parameter overlapped 1 for all of our
dipper function data sets (see top right panel in
Figure~\ref{fig-bayesianmodel}), indicating strong interocular
suppression regardless of post-retinal pathway. The flickering disc
stimuli produced a slightly lower suppression estimate, with
\(\omega < 0.9\). This may indicate slightly weaker suppression between
the eyes for temporal modulations, but it is much less extreme than our
recent estimates using EEG and matching paradigms (Segala et al., 2023).
One difference between these studies is the luminance of the background,
which was set to black for the experiments of Segala et al. (2023), but
here was set at the mean luminance. Other studies using static luminance
increments have also reported differences in the character of binocular
combination that are attributable to background luminance (Baker et al.,
2012), so this may explain the differences between studies. We also note
that the weaker interocular suppression for flickering disc stimuli
appears to be largely due to one participant (see
Figure~\ref{fig-individualdiscs}c), so individual differences might also
play a role. Future work should manipulate the background luminance
systematically to better understand how this modulates interocular
suppression.

@@

In V1, the classical view is that chromatic stimuli are processed in
`blob' regions that are largely monocular as they fall within ocular
dominance columns (Livingstone \& Hubel, 1984). However more recent work
has shown chromatic processing outside of the blobs (Sincich \& Horton,
2005), especially for stimuli with spatial structure (Chatterjee et al.,
2021). Since our psychophysical results indicate that interocular
suppression is equally strong within chromatic and achromatic pathways,
it may be that psychophysical performance indexes a common population of
(non-blob) neurons for all of our stimuli. On the other hand, if blob
regions include a subset of monocular neurons subject to interocular
suppression (e.g. Dougherty et al., 2019), this might be followed by
binocular summation at a later stage of processing for chromatic
stimuli, and potentially also for very low spatial frequency luminance
modulations also processed in blobs (Economides et al., 2011).

@@

\leavevmode\vadjust pre{\hypertarget{ref-Chatterjee2021}{}}%
Chatterjee, S., Ohki, K., \& Reid, R. C. (2021). Chromatic micromaps in
primary visual cortex. \emph{Nat Commun}, \emph{12}(1), 2315.
\url{https://doi.org/10.1038/s41467-021-22488-3}

@@

\leavevmode\vadjust pre{\hypertarget{ref-Sincich2005}{}}%
Sincich, L. C., \& Horton, J. C. (2005). The circuitry of V1 and V2:
Integration of color, form, and motion. \emph{Annu Rev Neurosci},
\emph{28}, 303--326.
\url{https://doi.org/10.1146/annurev.neuro.28.061604.135731}
